---
title: "DataAnalysis"
author: "Gabrielle McCabe"
date: "20/01/2020"
output: html_document
---

# Read and Manipulate Measurements Data

Import dplyr and ggplot2 and read measurements.csv.

```{r, echo = F}
library("dplyr")
library("ggplot2")

MData <- read.csv("Data/measurements.csv", header = TRUE)
```


Change the column names to shorter, 3-letter abbreviations. Spp: species. LbW: limb width. UtW: units for width. LbL: limb length. UtL: units for length. 

```{r}
MData <- rename(MData, Spp = Organism, LbW = Limb.Width, UtW = unitsW, 
                LbL = Limb.Length, UtL = unitsL)
```


Convert all limb width and limb length values to centimetres and save them in 2 new columns (LWc: limb length cm; LLc: limb length cm).

```{r}
MData$LWc <- MData$LbW
for(i in 1:length(MData$LWc)) {
  if(MData$UtW[i] == "mm") {
    MData$LWc[i] <- MData$LbW[i] / 10
  } 
}

MData$LLc <- MData$LbL 
for(i in 1:length(MData$LLc)) {
  if(MData$UtL[i] == "mm") {
    MData$LLc[i] <- MData$LbL[i] / 10
  } 
}
```


Create new column called Volume that calculates limb volume as: 
$$LimbVolume = pi * (width/2)^2 * length$$

```{r}
MData <- mutate(MData, Volume = pi * (LWc / 2) ^ 2 * LLc)
```


# Plots

Relationship between limb width (cm) and limb length (cm). Using the clean theme for presentations and publications used in the Colautti Lab.

```{r}
theme_pubworthy <- function (base_size = 12, base_family = "") {
  theme_classic(base_size = base_size, base_family = base_family) %+replace% 
    theme(
      axis.text = element_text(colour = "black"),
      axis.title.x = element_text(size=18),
      axis.text.x = element_text(size=12),
      axis.title.y = element_text(size=18,angle=90),
      axis.text.y = element_text(size=12),
      axis.ticks = element_blank(), 
      panel.background = element_rect(fill="white"),
      panel.border = element_blank(),
      plot.title=element_text(face="bold", size=24),
      legend.position="none"
    ) 
}

qplot(x = LWc, y = LLc, data = MData, size = I(5), alpha = I(0.6),
      xlab = "Limb Width (cm)", ylab = "Limb Length (cm)") + theme_pubworthy()
```


Relationship between limb width (cm) and limb length (cm) for species that have above average volume.

```{r}
Above.Avg.Spp <- filter(MData, Volume > mean(Volume))

qplot(x = LWc, y = LLc, data = Above.Avg.Spp, size = I(5),
      xlab = "Limb Width (cm)", ylab = "Limb Length (cm)") + theme_pubworthy()
```


# Table

First, I will filter out all the species and volume of species with below-average volume.

```{r}
Below.Avg.Spp <- MData %>% 
  filter(Volume < mean(Volume)) %>% 
  select(Spp, Volume)
Below.Avg.Spp
```


This table shows only species and volume for only the species with below-average volume.

Species | Volume
--- | ---
Felis catus | 11.28817
Mus musculus | 181.25198
Ateles fusciceps | 54.92002		
Bradypus tridactylus | 448.44593		
Pterois volitans | 83.73672		
Camelus dromedarius | 213.83794		
Ornithorhynchus anatinus | 1773.87773		
Pongo abelii | 183.27048	


# Grouping data

Add Habitat column. Set the seed (so the same random numbers are chosen) and randomly select 0 and 1 from binomial distribution then convert 0 to land and 1 to sea. 

```{r}
set.seed(3)
MData <- mutate(MData, Hab = rbinom(10, size = 1, prob = 0.5))

for(i in 1:length(MData$Hab)) {
  if(MData$Hab[i] == "0") {
    MData$Hab[i] <- "land"
    } else {
      MData$Hab[i] <- "sea"
      }
}
```


Summarize the mean and standard deviation of limb length (cm), limb width (cm), and volume for each of the habitat groups.

```{r}
MData %>% 
  group_by(Hab) %>% 
  summarise(Mean.Length = mean(LLc), SD.Length = sd(LLc), 
            Mean.Width = mean(LWc), SD.Width = sd(LWc),
            Mean.Volume = mean(Volume), SD.Volume = sd(Volume))
```

This table shows the mean and standard deviation (SD) of limb length (cm), limb width (cm), and volume for each of the habitat groups. 

Habitat | Mean Limb Length (cm) | SD Limb Length (cm) | Mean Limb Width (cm) | SD Limb Width (cm) | Mean Volume | SD Volume
--- | --- | --- | --- | --- | --- | ---
land | 105.6568 | 130.3758 | 4.920396 | 3.925232 | 5125.103 | 10205.79
sea | 112.6543 | 102.7764 | 7.224463 | 9.726004 | 19957.470 | 44256.16






