---
title: "Regex"
author: "Gabrielle McCabe"
date: "27/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


I imported dplyr and ggplot2 and read measurements.csv into MData. I included warning = FALSE and message = FALSE so that the warning messages from library("dplyr") did not show in the html.

```{r, warning = FALSE, message = FALSE}
library("dplyr")

MData <- read.csv("Data/measurements.csv", header = TRUE)
```


I added a new column called Sp that modified the Species names (in the Organism column) with only the first letter and a period. I modified the names by using the regular expression sub() to find and replace each species name in the Organism column with the first letter and a period, and then I used the dplyr command mutate() to add this column to MData. 

```{r}
MData <- mutate(MData, Sp = c(sub("Felis", "F.", Organism[1]), sub("Mus", "M.", Organism[2]), sub("Ateles", "A.", Organism[3]), sub("Bradypus", "B.", Organism[4]), sub("Pterois", "P.", Organism[5]), sub("Camelus", "C.", Organism[6]), sub("Ornithorhynchus", "O.", Organism[7]), sub("Canis", "C.", Organism[8]), sub("Pongo", "P.", Organism[9]), sub("Panthera", "P.", Organism[10])))

MData
```


I created a new dataset (spMData) containing only  the Sp column, moving it to the beginning in place of the original Organism column. I used the dplyr select() to rearrange MData, putting the species column first and then the remaining columns (Limb.Width, unitsW, Limb.Length, then unitsL), excluding Organism.

```{r}
spMData <- select(MData, Sp, Limb.Width : unitsL)
spMData
```














